Making multiple-locus association mapping on a genome-wide scale routine

### 2016-03-14 AM+draft.tex
150316 Working on results - sim - performance - elapse time

\documentclass[12pt]{article}
\linespread{1.5}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

@@

\subsection{Computer implementations for Multiple Locus Association Mapping}

We have chosen six computer implementations whose performance we  
compared  to our R package, am+. These are bigRR, glmnet, , lmm-lasso,
MLMM,  piMASS, and randfor.  We chose 
these six implementations because a. they were purpose-built for 
multiple-locus association mapping or contained general methods from 
which multiple-locus association mapping techniques had been developed, 
b. they could handle analysis on a genome wide scale, c.  they were freely 
available, and d. they are state-of-the-art.

A summary of the key 
features of the programs/packages whose 
performance we evaluated in this paper are given in table X.
Even from this small set of 
packages and computer programs, the diversity of implementation is apparent. 
BigRR and glmnet are based 
on linear models, lmm-lasso and MLMM on linear mixed models, piMASS on Bayesian models, 
and randfor on decision tree models. BigRR fit its models to data with ridge regression, a regularisation technique. 
Glmnet and lmm-lasso instead rely on lasso where parameter estimation and variable selection occur simultaneously. 
PiMass fits its models with bayesian variable selection, MLMM through stepwise selection, and randfor with random forests. 
Some of the implementations evaluated here fit all snp simultaneously where others do not. Although the underlying statistical 
problem was the same, to find the "best" set of predictions (snps) to explain a quantitative response (trait), the implementations differed greatly.

>>>>

\documentclass[12pt]{article}

\linespread{1.5}
\addtolength{\oddsidemargin}{-2.5cm}
%\addtolength{\evensidemargin}{-3cm}
\addtolength{\textwidth}{2.5cm}
\begin{document}

@@

\subsection{Computer programs for   multiple locus association mapping}

We compared the performance of our R package, am, to six other computer programs for multiple locus association mapping. These programs are ..., ..., ... We decided on these six because they implement published methods for multiple locus association mapping.  are capable of analysing association data on a genome wide scale, and are freely available. A summary of each of these programs along with am+ is given in table X. In preparing this table, we assumed the purpose of analysing data from a genome wide association study is to identify those snp in true association with a trait.  From table X, the computer programs vary greatly in methodology, implementation, requirements, and utility.  For example, some programs have methodology that fits all marker-trait associations simultaneously where others  fit only a subset of marker-trait associations jointly. Some programs have tuning parameters where others have none. Several cannot handle additional fixed effects such as covariates to account for population structure. Also, the majority of programs considered here yield results that require additional computation before a "best" set of snp is realised. Only am+  fits all marker trait associations simultaneously, handles different input data, accommodates additional fixed effects, has no parameters to be fine tuned, returns the "best" set of snp in joint association with a trait without additional computation, and is easy to use.

\subsection{Computational performance}

To assess the computational performance of am+, we conducted a simulation study. Our performance measures are computing time and memory usage. We also compared the performance of am+ to six other computer programs for multiple locus association mapping (table X).

Data were generated under five different scenarios. These scenarios are motivated by the type and size of gwas currently being performed. The five scenarios are  unsequenced crop plants (UCP, 150 plants and 5k snp), model plants ( 4555 plants and 34544k snp), ..... Our marker data are from the 1000 Genomes project (Phase 3). Quantitative trait data are generated from these data by assigning effect sizes to a randomly selected set of snp loci along with a gaussian deviate for the error. 
The size of the error is  set to yield a heritability of 0.5.
